### 2023-04-22 chapter2/chapter2.tex
Moved substitution figure between results paragraphs
Added introduction outline
Updated WIP analysis of scores outline
Added description of GO term analysis to results
Converted outline to paragraph and tweaked wording

\section{Introduction}
Here's where my introduction will go!

@@

% RESULTS
% III. Brownian motion score analyses
%     A. Motivation for use of phylogenetic models
%         1. For each residue in the input sequence, AUCPreD returns a score between 0 and 1 where higher values indicate higher confidence in a prediction of structural disorder.
%     B. When examining the disorder score traces in greater detail alongside the alignments, we were struck by the significant variability in some traces, even when there was a relatively high level of sequence identity.
%     C. To better understand the sequence features underlying these changes in predicted disorder, we calculated the values for 82 disorder-associated features in each region
%     D. Though this analysis is limited by the resolution of the regions, which varies from 30 to ??? residues, it indicates the prediction algorithm is sensitive to these features in bulk.
%          1. Some more observations of the overall trends ...
%     E. Interestingly, the correlations with many motifs were statistically significant, though small in magnitude relative to the non-motif features.
%          1. Look into motif regexes and see if ones with highest correlations have any common features
%     F. To determine if regions with rapidly evolving disorder scores were enriched in any processes, we extracted the upper decile of their rate distribution
%          1. Many chromatin and DNA related features

% IV. Brownian motion feature analyses
%     A. Having calculated features for all IDRs in our data set, we were curious the distribution contained any global structure which would allow the identification of classes with distinct biophysical or functional properties
%     A. Feature roots
%         1. Discuss differences in scaling and need for z-scores
%         2. Discussion of sample regions in figure? (need to add)
%     B. The phylogenetic contrasts algorithm also permits an estimate of the rate of feature evolution, which is effectively the variance of the distribution of features across all sequences after correcting for the phylogenetic relationships between them
%         1. PC1 correlates with rate
%         2. PC2 and PC3 show three major axes of unconstrained variation: Q-rich, G-rich, acidic

% V. BM-OU analyses
%    A. While these analyses measure the rates of change of features after accounting the the phylogenetic relationships between sequences, they do not directly test for conservation of features
%         1. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time
%         2. Instead, demonstrating evidence of conservation requires a comparison to a model where feature evolution is constrained.
%         3. A common choice is the Ornstein-Uhlenbeck (OU) model
%             i. Some explanation about how it works

>>>>

\section{Introduction}
Here's where my introduction will go!
% I. IDRs, proteins without defined structures, are prevalent in eukaryotic proteins and play key roles in many processes
% II. Sequence-function relationships are unclear in IDRs
% III. Proteome-wide patterns of conserved features linked to function
%    A. Past studies have demonstrated evidence of constraint in specific proteins for specific properties
%    B. Additionally other studies have shown importance of short linear motifs in IDRs
%        1. Can be identified as short segments of conservation in otherwise highly diverged alignments of IDRs
%        2. However, these segments comprise a small fraction of disordered regions, so
%    C. Recent study first to demonstrate patterns of selection called evolutionary signatures in IDRs across entire proteome using a set of 82 molecular features
%    D. Furthermore, by clustering IDRs on by these evolutionary signatures, they identified groups of IDRs evolving under similar sets of constraints. Furthermore they identified functional associations in these clusters which can be used to generate hypotheses of function for previously uncharacterized proteins
% IV. Unclear if these patterns are specific to yeast or representative of IDR evolution across the tree of life
%    A. Expand analyses to another model organism
%        1. Drosophila is key model for genetic and development research with abundant genomic resources across many lineages in the radiation
%        2. Furthermore as multicellular organism with complex developmental pathways, the results would be particularly relevant for understanding the role of IDRs in complex signaling pathways
%        3. Also relevant for human health!
%    B. Results would have significant impact on interpretation
%        1. Conserved distributed properties are a yeast phenomena
%        2. Patterns of conservation are yeast phenomena
%        3. Global taxonomy of disordered regions similar to Pfam
% IV. Explicit phylogenetic models of trait evolution can reveal patterns of evolution with fewer assumptions
%    A. Previous analyses compared observed distributions of features in observed IDRs to the distributions generated by simulated counterparts
%    B. Differences were interpreted as evidence of selection on features
%        1. However, evidence of deviation from the null model rather than evidence of constraint necessarily
%        2. Null distribution generated by a model of evolution where the composition of sequences stochastically evolve towards a specific equilibrium distribution
%        3. Previous studies have identified IDRs with varying compositional preferences, so a single equilibrium distribution is unlikely to globally describe observed patterns of sequence evolution in IDRs
%    C. Thus in this work we explore use of various statistical approaches that directly measure rates of change and constraint of molecular features indepedent of specific model of sequence evolution
%        1. Place these evolutionary analyses in a statistical phylogenetic framework
%        2. Though these models have own assumptions, they are macroscopic descriptions of continuous trait evolution and therefore independent of any assumptions about evolution at the level sequences
%        3. Still have their own assumptions of course, but more general and robust to deviations
%        4. We view these approaches as complementary, each having different trade-offs between specificity and generality of testing specific hypotheses of disorder evolution
%        5. We also explore hybrid approaches where the evolutionary rates obtained from phylogenetic models are used as a statistic
% V. By applying these approaches to set of IDRs identified in Drosophila genomes, we observe...

@@

\subsection{Intrinsic disorder is poorly conserved in some proteins}
Though the substitution models reveal specific patterns of evolution at the level of individual residues, the large amounts of sequence divergence between many orthologous IDRs implies their evolution is not well-described by fine-scale models of residue substitution. Given the growing evidence that IDRs are constrained to conserve distributed properties, we instead turned towards characterizing their evolution in terms of 82 disorder-associated ``molecular features'' obtained from the previous study of IDRs in the yeast proteome. However, before conducting an in-depth analysis of these features, we examined the disorder score traces in greater detail and were struck by the significant variability between species. For each residue in the input sequence, AUCPreD returns a score between 0 and 1 where higher values indicate higher confidence in a prediction of intrinsic disorder. In some alignments, the scores would vary by nearly this entire range at a given alignment position even when there was a relatively high level of sequence identity (Fig.~\ref{fig:scores}A).

To better understand the relationship of this variability to differences in the regions' biophysical properties, we sought to correlate the average disorder score of the segments in a region with their molecular features. However, because the sequences are evolutionarily related, they are not independent, making them unsuitable for direct use in many standard statistical procedures. We instead applied the method of contrasts, which takes differences between adjacent nodes in the phylogenetic tree relating the species to generate a set ``contrasts,'' to both the scores and the features. Under some general assumptions of the underlying evolutionary process, contrasts are independent and identically-distributed and are therefore appropriate for use in correlation analyses. The resulting feature contrasts have varying degrees of correlation with the score contrasts (Fig.~\ref{fig:scores}B-C). Some, like isopoint, are uncorrelated, but most are significantly, if weakly, correlated. In general, the strongest correlations are observed for features which have a direct biophysical relationship to the presence or absence of disorder, such as disorder\_fraction or hydrophobicity. Interestingly, the correlations with many motifs were statistically significant, though small in magnitude relative to the non-motif features. However, a more detailed analysis of this observation is presented in the discussion. To determine if regions with rapidly evolving disorder scores were associated with a particular function, process, or compartment, we then extracted the upper decile of their rate distribution and performed term enrichment analyses on their associated annotations (Fig.~\ref{sfig:scores_histogram}). The most significant terms are generally related to DNA repair or extracellular structure, which suggests these processes and components are enriched in proteins whose structural state is rapidly evolving (Fig.~\ref{fig:scores}E).

% RESULTS
% IV. Brownian motion feature analyses
%     A. Having calculated features for all IDRs in our data set, we were curious the distribution contained any global structure which would allow the identification of classes with distinct biophysical or functional properties
%     A. Feature roots
%         1. Discuss differences in scaling and need for z-scores
%         2. Discussion of sample regions in figure? (need to add)
%     B. The phylogenetic contrasts algorithm also permits an estimate of the rate of feature evolution, which is effectively the variance of the distribution of features across all sequences after correcting for the phylogenetic relationships between them
%         1. PC1 correlates with rate
%         2. PC2 and PC3 show three major axes of unconstrained variation: Q-rich, G-rich, acidic

% V. BM-OU analyses
%    A. While the the BM model can the rates of change of features after accounting the the phylogenetic relationships between sequences, they do not directly test for conservation of features
%         1. In fact, under the BM model, trait variation is unconstrained and will increase without bound over time
%         2. Instead, demonstrating evidence of conservation requires a comparison to a model where feature evolution is constrained.
%         3. A common choice is the Ornstein-Uhlenbeck (OU) model
%             i. Some explanation about how it works
